\documentclass[]{article}

\begin{document}

We would like to thank the referee for the positive report.
We have taken the comments into account in the following way:

\begin{itemize}
\item  Abstract: we have replaced ``results for 14 TeV are presented" by ``results for the LHC at 14 TeV are presented".

\item Introduction: we have replaced\\
 ``increase with larger centre-of-mass energy (e.g. $\sqrt{s}=27$ or 100\,TeV)" by \\
``increase with larger centre-of-mass energy (e.g. $\sqrt{s}=27$ (HE-LHC) or 100\,TeV (FCC-hh))".

\item We have added the following sentence in the conclusions:
``We have demonstrated that the dependence of both the total and the
differential K-factors on the value of $\lambda$ is stronger than the $m_t\to\infty$ limit may suggest."

\item We have added the scale uncertainties for the 27\,TeV case in Table 1.

\item References: 
\begin{itemize}
\item 1811.07906 (Dreyer, Karlberg): 
Our intention was to focus on $HH$ production in gluon fusion, otherwise citations to a plethora of
$HHjj$ (VBF and non-VBF), $HHZ,HHW, HHt\bar{t}$ calculations would have to be included.  
Therefore we would prefer to restrict the references to the ones
related to $HH$ production in gluon fusion.
The analogous is done in 1811.07906, as far as we can see (even though
some $gg\to HH$ calculations are mentioned, strange enough in the context of
single Higgs production). 

\item We have included the references Li et al [arXiv:1904.12006] and Corbett et al [arXiv:1705.02551] in the list of phenomenological studies.

\item We have included Davies et al [arXiv:1904.11998] when mentioning the high energy expansion.
\end{itemize}

\end{itemize}

\end{document}
