We have presented results for Higgs boson pair production in gluon
fusion at full NLO for non-standard values of the trilinear Higgs
boson coupling $\lambda$. We have also shown how results with modified 
top quark Yukawa coupling can be produced with the same code.
The total cross section is a quadratic
polynomial in $\lambda$, with a minimum around $\chhh\approx 2.4$,
which is present both at LO and NLO with full top quark mass
dependence, stemming from destructive interference of diagrams with
and without a trilinear Higgs coupling. 
The $\mhh$ distribution shows a dip around this minimum, which is to
less extent also visible in the transverse momentum distribution of
one of the Higgs bosons. 
We have assumed in our study that modifications of the Higgs couplings
to other particles are small and can be increasingly well constrained by
other processes. Nonetheless it should be kept in mind that a dip in
the $\mhh$ distribution could also originate from other effective
couplings, for example an effective $t\bar{t}HH$ coupling, while
$\chhh= 1$~\cite{Buchalla:2018yce}.

We have also combined our NLO QCD results with the \pythia and \herwig
parton showers, where in the \herwig case we employed both the default 
shower (the angular-ordered $\tilde{q}$-shower) and the dipole shower.
We observed that for distributions particularly sensitive to the 
additional radiation, the parton showers exhibit a somewhat different 
behaviour. While both \herw showers generate comparable results and 
perform as expected in the NLO regime, the \pythia shower produces harder 
radiation, for example in the tail of the $\pthh$
distribution. Varying the shower hard scale in \herwig on top of
$\mu_R,\mu_F$ variations leads to
uncertainty bands which approximately cover these 
differences. However the parton shower uncertainties 
then can become sizeable and even surpass the fixed-order scale uncertainties.

The \powheg{} version of the code for Higgs boson pair production
including the possibility to vary the trilinear coupling 
and the top quark Yukawa coupling 
is publicly available in the \powhegbox{\tt-V2} package at the website
{\tt http://powhegbox.mib.infn.it}, in the
{\tt User-Processes-V2/ggHH/} directory.
